---
title: "eli review"
output:
  pdf_document: default
  html_document: default
date: "2024-05-25"
---
##2a) Analysis of my own dataset

This dataset involves analyzing the number of shootings in NYC over time from January 2006 to December 2022. The data is taken from NYC's public data repository. I had to clean the data because the data before only gave me a table of 27,000 values of the date, time, location, etc. of the shooting. Thus the dataset I'm using is a self-modified version of the original one.
```{r running SARIMA/ARIMA}
library(astsa) #acf
library(readr)
library(tseries) #ADF test
library(forecast) #auto.arima function

#Reading the data into R as a time series
dict <- read_csv("dict.csv")$Shootings 
shootings_ts <- ts((dict), start=c(2006, 1), end=c(2023, 12), frequency=12) 

#Plotting our shooting data as a time series
plot(shootings_ts, xlab='Year', ylab='Number of Shootings', main='Monthly Shootings in NYC Jan 2006 - Dec 2023', col='steelblue1', xaxt='n')
years <- seq(2006, 2023, by=1)
for (year in years) {
  abline(v=year, col="lightgray", lty=2)
}
axis(1, at=seq(2006, 2023, by=1), labels=seq(2006, 2023, by=1), cex.axis=0.7) 

#Analysis of the data
adf.test(shootings_ts) # adf gives result as stationary
qqnorm(shootings_ts, main='QQ plot of Raw Data')
qqline(shootings_ts) # the curve in the points indicates that data is not normal
acf2(shootings_ts) # acf and pacf show heavy seasonality

#Since the data is non-normal, we apply Box-Cox transformation
shootings_ts1<-diff(log(shootings_ts))
ts.plot(shootings_ts1, main="Differenced 1 Monthly Shootings in NYC Jan 2006 - Dec 2023")

qqnorm(shootings_ts1, main='QQ Plot of Difference 1') #This plot looks much better, basically normal/stationary
qqline(shootings_ts1)
acf2(shootings_ts1, 12)
# Since this doesn't give us a simple model for our SARIMA, we run auto.arima

# Predictions
shootings_modeled<-auto.arima(shootings_ts)
checkresiduals(shootings_modeled) # statistically checking the model
summary(shootings_modeled) # the residuals look normal 

# Forecasting future 12 values
future <- forecast(shootings_modeled, h = 12)
print(future)

plot(future, main='NYC Shootings Forecasted Values n=12', xlab = 'Year', ylab='Number of Shootings', xaxt='n')
axis(1, at=seq(2006, 2024, by=1), labels=seq(2006, 2024, by=1), cex.axis=0.7) 


```

```{r 3.2 and 4.2 Spectral Analysis}
weather_dict <- read_csv("new york.csv")$Value # I pulled in New York Weather Data from Jan 2006 - Dec 2023
weather_ts <- ts((weather_dict), start=c(2006, 1), end=c(2024, 5), frequency=12) 

sr = mvspec(cbind(weather_dict, dict), kernel("daniell",2), plot=FALSE)
sr$df
f = qf(.95, 2, sr$df-2)
C = f/(18+f)
plot(sr, plot.type = "coh", ci.lty = 2, main='Coherence Between Weather and Shootings')
abline(h = C, col='pink')

periodogram <- spec.pgram(shootings_ts1, plot=FALSE)
plot(periodogram, log = "y", main = "Periodogram of Shootings") # Frequencies are especially strong every cycle

x.spec <- spectrum(shootings_ts, main= "Spectral Analysis of NYC Shootings", log="no")
x.logdif1 <- spectrum(shootings_ts1, main= "Log Spectral Analysis of NYC Shootings", log="no") # The transformed data provides clearer signals. We can see the dominant frequency occurs after one cycle. Additionally, there is a weaker peak at 5 cycles- which could be some other factor


```

